---
title: "Eating Out"
author: "Sisilia Martinez"
output: 
  html_document:
    number_sections: true
    toc: true
    toc_depth: 5
date: "`r Sys.Date()`"    
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The data comes from the 1998 General Social Survey answer this question

$H_0$ There is no difference among the various races with frequency of eating out  
$H_A$ There is a statistical difference between at least two races with the frequency of eating out  


```{r include=FALSE,echo=FALSE}
source(file="fisherXY.R")
require(tidyverse)
require(tigerstats)
require(knitr)
require(foreign)
GSS <-read.dta(file="GSS1998.dta")
GSS <- droplevels(GSS)
mydata <- select(GSS,race,age,eatout)
mydata <- na.omit(mydata)
```


```{r}
barchartGC(~race + eatout,data=mydata)
barchartGC(~race+ eatout,data=mydata, type="percent")

```

Looking at the bar chart you can infer that the white race eat out more than the other races.

```{r}
table1 <- xtabs(~race + eatout, data=mydata)
rowPerc(table1)
colPerc(table1)
```

Looking at the numerical summary it tells us that for the white race has 83.11% eat out and 16.89% do not. For the black race 69.04% eat out and 30.96% do not. Lastly for other races 72.12% eat out and 27.88% do not. You can see that white people and other races eat out more compared to black people.

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```


